Water desalination using polyelectrolyte hydrogel. Gibbs ensemble modeling

### 2022-07-19 gibbs.tex
TOC figure

\affiliation{Department of Physical and Macromolecular Chemistry, Faculty of Science,
Charles University in Prague, Czech Republic}

\alsoaffiliation[imc]{Institute of Macromolecular Compounds of Russian Academy of Sciences,
Saint-Petersburg, Russia}

@@

\begin{figure*}[t]
\centering \includegraphics[width=0.7\textwidth]{figures/simgibbs}\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:diamond}.}
\end{figure*}

@@

\begin{figure*}[t]
\centering 
\subfloat[Open system \label{fig: open}]
{\includegraphics[width=0.5\textwidth]{figures/gc}}
\subfloat[Closed system \label{fig: closed}]
{\includegraphics[width=0.5\textwidth]{figures/gibbs}}
\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:open and closed}.}
\end{figure*}
We simulate the compression of the gel in two ensembles.
\begin{enumerate}
\item The first is so called \emph{open system}, when the simulation box with gel freely exchanges ions with a big amount of aqueous solution of certain constant salinity, $\cs$. 
\item And in \emph{closed system}, when the gel is in equilibrium with a finite volume of aqueous solution.
\end{enumerate}
The compression of the gel in the \emph{open} system does not affect the surrounding salinity, thus the density of ions outside the gel remains constant $\cs =$ $\cna=$ $\ccl=$ $\text{Const}$ (\reffig{fig: open}). 
On the contrary, what remains constant in the \emph{closed} system, is the total number of $\na$ and $\cl$ ions, that are contained in the gel and in the external volume, \ie in the volume $\Vbox$ (\reffig{fig: closed}).

@@

\paragraph{Compression in open system.}
\begin{figure*}[t]
\subfloat[The gel partial pressure vs gel molar volume \label{fig: PV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_PV_pub}}
\hspace{0.02\textwidth}
\subfloat[Supernate salinity vs gel molar volume\label{fig: CV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_CV_pub}}
\caption{
The compression of the gel in open system (solid lines) and in closed system (dotted lines). 
Each solid curve corresponds to different salinity of the reservoir $\cs$ (see legend). Shadowed area limit the states with applied pressure below zero and above 5 bar.\label{fig: PV and CV}}
\end{figure*}
First, we run a set of simulations modeling the gel compression in \emph{open system}, \ie in equilibrium with a big bath of certain salinity, $\cs$. 
The simulations are run for a set of different simulation box volumes. 
Each simulation returns the averaged pressure, $P$, and the number of $\cl$ ions, which are present in the simulation box, $\ncl$.

@@

\begin{figure*}[t]
\subfloat[Pressure vs gel molar volume \label{fig: NV}]
{\includegraphics[width=0.49\textwidth]{figures/fig_NV_pub}}
\subfloat[Salinity vs gel molar volume\label{fig: CN}]
{\includegraphics[width=0.495\textwidth]{figures/fig_NVbox_mu_pub}}
\caption{The compression of the gel in \emph{open system} (solid lines) and in \emph{closed system} (dotted lines). 
Shadowed area limit the states with applied pressure below zero and above 5 bar.
The values $N_\cl$ are the virtual numbers of present $\cl$ ions in \emph{closed system} simulation boxes.
\label{fig: NV and CN}}

@@

\section{Conclusions}
We have modeled the compression of the gel in thermodynamic equilibrium with the reservoir of limited amount of water.
We have shown that in case of gel made of strong polyelectrolyte, the compression always lead to decrease of the surrounding salinity.
We have modeled the process of water desalination combining two processes: 
(1) the swelling of the gel in \emph{open} system, exchanging ions with big reservoir at constant salinity. 
(2) the compression of the gel in \emph{closed} system, when the gel affect the salinity in surrounding solution.
We estimated the energy needed to produce one liter of water and have show that the method 
%at least theoretically 
may compete with the modern technologies used in modern industry.

@@

This research was supported by the Czech Science Foundation (grant
19-17847Y) (OR, LN, AK), Government of Russian Federation, grant number
14.W03.31.0022. AK thanks the Grant Agency of Charles University for
support (project 318120). Computational resources were supplied by
the project ``e-Infrastruktura CZ'' (e-INFRA LM2018140) provided
within the program Projects of Large Research, Development and Innovations
Infrastructures.

>>>>

\affiliation{Department of Physical and Macromolecular Chemistry, Faculty of Science, Charles University in Prague, Czech Republic}
\alsoaffiliation[imc]{Institute of Macromolecular Compounds of Russian Academy of Sciences,
Saint-Petersburg, Russia}

@@

\begin{figure*}[t]
\centering \includegraphics[width=0.7\textwidth]{{figures/simgibbs}.pdf}\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:diamond}.}
\end{figure*}

@@

\begin{figure*}[t]
\centering 
\subfloat[Open system \label{fig: open}]
{\includegraphics[width=0.5\textwidth]{figures/gc.pdf}}
\subfloat[Closed system \label{fig: closed}]
{\includegraphics[width=0.5\textwidth]{figures/gibbs.pdf}}
\caption{Diamond-like network in the simulation box. Color code represents
the individual ion types (red: $\na$, blue: $\cl$, yellow: $\ca$)
and the hydrogel (gray: neutral segment ($\AH$), cyan: charged segment
($\A$)). \label{fig:open and closed}.}
\end{figure*}
We simulate the compression of the gel in two ensembles.
\begin{enumerate}
\item The first is so called \emph{open system}, when the simulation box with gel freely exchanges ions with a big amount of aqueous solution of certain constant salinity, $\cs$. 
\item And in \emph{closed system}, when the gel is in equilibrium with a finite volume of aqueous solution.
\end{enumerate}
The compression of the gel in the \emph{open} system does not affect the surrounding salinity, thus the density of ions outside the gel remains constant $\cs =$ $\cna=$ $\ccl=$ $\text{Const}$ (\reffig{fig: open}). 
On the contrary, what remains constant in the \emph{closed} system, is the total number of $\na$ and $\cl$ ions, that are contained in the gel and in the external volume, \ie in the volume $\Vbox$ (\reffig{fig: closed}).

@@

\paragraph{Compression in open system.}
\begin{figure*}[t]
\subfloat[The gel partial pressure vs gel molar volume \label{fig: PV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_PV_pub.pdf}}
\hspace{0.02\textwidth}
\subfloat[Supernate salinity vs gel molar volume\label{fig: CV}]
{\includegraphics[width=0.48\textwidth]{figures/fig_CV_pub.pdf}}
\caption{
The compression of the gel in open system (solid lines) and in closed system (dotted lines). 
Each solid curve corresponds to different salinity of the reservoir $\cs$ (see legend). Shadowed area limit the states with applied pressure below zero and above 5 bar.\label{fig: PV and CV}}
\end{figure*}
First, we run a set of simulations modeling the gel compression in \emph{open system}, \ie in equilibrium with a big bath of certain salinity, $\cs$. 
The simulations are run for a set of different simulation box volumes. 
Each simulation returns the averaged pressure, $P$, and the number of $\cl$ ions, which are present in the simulation box, $\ncl$.

@@

\begin{figure*}[t]
\subfloat[Pressure vs gel molar volume \label{fig: NV}]
{\includegraphics[width=0.49\textwidth]{figures/fig_NV_pub.pdf}}
\subfloat[Salinity vs gel molar volume\label{fig: CN}]
{\includegraphics[width=0.495\textwidth]{figures/fig_NVbox_mu_pub.pdf}}
\caption{The compression of the gel in \emph{open system} (solid lines) and in \emph{closed system} (dotted lines). 
Shadowed area limit the states with applied pressure below zero and above 5 bar.
The values $N_\cl$ are the virtual numbers of present $\cl$ ions in \emph{closed system} simulation boxes.
\label{fig: NV and CN}}

@@

\section{Conclusions}
We have modeled the compression of the gel in thermodynamic equilibrium with the reservoir of the limited amount of water.
We have shown that in the case of gel made of strong polyelectrolyte, the compression always decreases the surrounding salinity.
We have modeled the process of water desalination by combining two processes: 
(1) the swelling of the gel in \emph{open} system, exchanging ions with a big reservoir at constant salinity. 
(2) the compression of the gel in \emph{closed} system, when the gel affects the salinity in the surrounding solution.
We estimated the energy needed to produce one liter of water and have shown that the method may compete with the modern technologies used in modern industry.

@@

This research was supported by the Czech Science Foundation (grant 19-17847Y)
Government of Russian Federation, grant number 14.W03.31.0022. 
Computational resources were supplied by the project ``e-Infrastruktura CZ'' (e-INFRA LM2018140) provided within the program Projects of Large Research, Development and Innovations Infrastructures.
